Issues with Dynamical Measurements Across the Disk

### 2019-02-11 paper/ms.tex
Change title
Tweaks to abstract

\title{Our Galactic Midplane Is Probably Wrong, and Not by a Little}

@@

We propose a new type of error in the computation of stellar actions. This
error is induced by the fact that the local Galactic midplane varies across
the disk. As a result, when we make 6D phase space measurements of stars that
are non-local (i.e. far away), we are observing those stars with a systematic
offset in $z$. We show that this systematic offset induces time evolution in
the observed actions which we interpret as an error. We note that this error
is non-Gaussian and bimodal, with neither of the modes peaking on the true
action value. A midplane error of $\sim30\,\pc$ and $\sim225\,\pc$ is
necessary to induce a $100\,\%$ error in the vertical action $J_z$ for thin
and thick disk orbits, respectively. Using cosmological zoom-in simulations
from the FIRE collaboration, we show that the midplane can vary by
$\sim100\,\pc$ around the Solar circle, indicating that such an effect is
important for understanding the thin and thick disk. Our discussion is also
directly applicable to variations in the local standard of rest. The variation
of the midplane must be taken into account when performing dynamic modelling
across large regions of the disk accessible with the {\em Gaia} mission.

>>>>

\title{Issues with Dynamical Measurements Across the Disk}

@@

We demonstrate the effect of midplane errors on the computation of stellar
actions. This error is induced by the fact that the local Galactic midplane
varies across the disk. As a result, when we make 6D phase space measurements
of stars that are non-local (i.e. far away), we are observing those stars with
a systematic offset in $z$. We show that this systematic offset induces time
evolution in the observed actions which we interpret as an error. We note that
this error is non-Gaussian and bimodal, with neither of the modes peaking on
the true action value. A midplane error of $\sim30\,\pc$ and $\sim225\,\pc$ is
necessary to induce a $100\,\%$ error in the vertical action $J_z$ for thin
and thick disk orbits, respectively. Using cosmological zoom-in simulations
from the FIRE collaboration, we show that the midplane can vary by
$\sim100\,\pc$ around the Solar circle, indicating that such an effect is
important for understanding the thin and thick disk. Our discussion is also
directly applicable to variations in the local standard of rest. The variation
of the midplane must be taken into account when performing dynamical modelling
across large regions of the disk accessible with the {\em Gaia} mission.
